---
title: "Loading the data"
author: "Me"
date: "2023-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

all libraries used in this document:

```{r}
library(readxl)
library(dplyr)
```

## Loading the dengue-excel-sheets and transforming them into dataframes

```{r}
Endung = "_dengue_extracted.xlsx"
Anfang = "Dengue"

for (i in 2006:2020) {
  
  Dateiname = paste0(i, Endung) #creating the file name
  Path = paste0("Excel-Data/", Dateiname)
  Name_df = paste0(Anfang, i) 
  assign(Name_df, read_excel(Path)) #creating the variable and assigning the excel sheet
  
  df_dengue = get(Name_df)
  vec_data_names = vector("character")
  
  vec_data_names = sapply(df_dengue$Reporting_areas, function(g) {
    paste(g[1], i)
  })
  data_names = matrix(vec_data_names,nrow = 77, ncol = 1)
  df_dengue = cbind(data_names, df_dengue)
  assign(Name_df,df_dengue)
}

#combinign the single years into one dataframe

total_years = data.frame()
for (i in 2006:2020){
  a = get(paste0("Dengue",i))
  total_years = rbind(total_years, a)
}
```

## Rows as time-scale

sorting the dataframe by province names transforming the dataframe, to move the time scale from the x-axis to the y-axis

```{r}
m = 1
data_total = data.frame(col1 = character(0), col2 = character(0), col3 = numeric(0), month = character(0))
colnames(data_total) = c("data_names", "Reporting_areas", "dengue_cases", "month")
sorted_total_years = total_years[order(total_years$"Reporting_areas"),]

for (k in 1:1155){
  
  for (j in 3:14){
    square = sorted_total_years[k, c(1, 2, j)] 
    month = colnames(square[3])
    square = cbind(square,month)
    colnames(square) = c("data_names", "Reporting_areas", "dengue_cases", "month")
    data_total = rbind(data_total, square)
    colnames(data_total) = c("data_names", "Reporting_areas", "dengue_cases", "month")
    data_total[m,1] = paste(data_total[m,1], month[1])
    m = m + 1
  }
}
```

## Adding temperature, longitude and latitude column

```{r}
m = 1
temp_data = read_excel("era5_data_2006_2020_thailand_monmean.xlsx")
temp_data = as.data.frame(temp_data)
temp_data_sorted = data.frame( col1 = numeric(0), Longitude = numeric(0), Latitude = numeric(0))
for (i in 1:77) {
  
  for (j in 5:184) {
    # sorting the temperature data to fit the format of data total
    temp_data_sorted[m,1] = temp_data[i,j]    
    temp_data_sorted[m,2] = temp_data[i,3]
    temp_data_sorted[m,3] = temp_data[i,4]
    m = m + 1
  }
}
data_total = cbind(data_total, temp_data_sorted) 
colnames(data_total)[5] = "temperature"
```

## Loading population data

```{r}
Endung = "_population.xlsx"
# empty dataframe for the data
total_population = data.frame(Reporting_areas = character(), population_count = numeric())

# stack years of 2006-2011 below each other
for (i in 2006:2011) {
  population_df  = data.frame()
  Dateiname = paste0(i, Endung) # creating the file name
  population_df = read_excel(Dateiname) # creating the variable and assigning the excel-sheet
  vec_years = rep(i, times = 77)
  population_df = cbind(population_df,vec_years)
  names(population_df) = c("Reporting_areas", "population_count", "year")
  # duplicate every row 12 times to get population for each month, because population is counted yearly
  repeat_count = 12
  population_df_replicated = population_df[rep(row.names(population_df), each = repeat_count), ]
  row.names(population_df_replicated) = NULL
  
  # put all the years together
  total_population = rbind(total_population, population_df_replicated)
  
}

# stack years of 2012-2020 below each other and add to data_total
population_df = read_excel("2012-2020_population.xlsx")
colnames(population_df) = c("Reporting_areas", "population_count", "population_count", "population_count", "population_count", "population_count","population_count","population_count","population_count","population_count")
pop_placeholder = data.frame(Reporting_areas = character(), population_count = numeric(), year = numeric())

col_num = 2
for (i in 2012:2020) {
  year = rep(i, times = 77)
  pop_placeholder = rbind(pop_placeholder, cbind(population_df[,c(1,col_num)],year))
  names(pop_placeholder) = c("Reporting_areas", "population_count", "year")
  col_num = col_num + 1
}

# duplicate every row 12 times to get population for each month, because population is counted yearly
repeat_count = 12
population_df_replicated = pop_placeholder[rep(row.names(pop_placeholder), each = repeat_count), ]
row.names(population_df_replicated) = NULL

# combining all population data in one dataframe and adding it to data total
total_population = rbind(total_population, population_df_replicated)
sorted_total_population = total_population[order(total_population$"Reporting_areas"),]

data_total = cbind(data_total,sorted_total_population[,c(2,3)])
row.names(data_total) = NULL
```

## Adding time column

time column gives an overview over year and month

```{r}
first_date = as.Date("2006-01-01")
last_date = as.Date("2020-12-31")

time_column_short = seq(first_date, last_date, by = "month")
time_column_short = as.Date(time_column_short, format = "%Y-%m")
time_column = rep(time_column_short, times = 77)

data_total = cbind(data_total, time_column)
#data_total = subset(data_total, select = -time_column) 
```

## Creating an incidence-column

makes infections comparable by normalising them according to population size

```{r}
# checking that all objects have the right data-type
data_total = as.data.frame(data_total)    
data_total$dengue_cases = as.integer(data_total$dengue_cases)
data_total$population_count = as.integer(data_total$population_count)
dimensions = dim(data_total)

R= dimensions[1]
for(Zeile in 1:R){
  Population = data_total$population_count[Zeile]
  Infection = data_total$dengue_cases[Zeile]
  Incidence = (Infection/Population)*100000
  data_total$incidence[Zeile] = Incidence
}
```

## Adding the data from Bungkan-district to Nong Khai and deleting all Bungkan-Data

Bungkan was only formed in 2011 from a part of Nong Khai, leading to missing values für the first 5 years in Bungkan. By adding the values together, the province of Nong Khai before the seperation remains the reporting area.

```{r}
for (i in 1:dim(data_total)[1]) { 
  
  if(data_total$Reporting_areas[i]=="Bungkan"){     
    
    Datum = data_total$time_column[i]
    for(j in 1:dim(data_total)[1]){
      if(data_total$Reporting_areas[j] == "Nong Khai" & data_total$time_column[j] == Datum){
        # overwriting NAs or adding the values together
        data_total$dengue_cases[j]= ifelse(is.na(data_total$dengue_cases[i]), data_total$dengue_cases[j], data_total$dengue_cases[i]+data_total$dengue_cases[j])
        data_total$population_count[j]= ifelse(is.na(data_total$population_count[i]), data_total$population_count[j], data_total$population_count[i]+data_total$population_count[j])
        data_total$incidence[j]= ifelse(is.na(data_total$incidence[i]), data_total$incidence[j], data_total$incidence[i]+data_total$incidence[j])
        
      }
    }
  }
}
```

## Creating clean dataframe

detecting all NA and removing the rows containing them, creating a cleaned dataframe to use when NAs would lead to errors

```{r}
detect_na = function(x) {
  is_na = x %in% c("N/A", "na", "NA", "n/a")
  replace(x, is_na, NA)
}
data_total = apply(data_total, 2, detect_na)
rmv.rows = apply(data_total,1,function(x){sum(is.na(x))})
i.missing = which(rmv.rows >0)

clean_data_total = as.data.frame(data_total[-i.missing,])
num_col = c("temperature", "Longitude", "Latitude", "incidence")
int_col = c("dengue_cases","population_count","year")
date_col = c("time_column")

clean_data_total[int_col] = lapply(clean_data_total[int_col], as.integer)
clean_data_total[num_col] = lapply(clean_data_total[num_col], as.numeric)
clean_data_total[date_col] = lapply(clean_data_total[date_col], as.Date)
summary(clean_data_total)
```

## Structuring the dataframe and converting into the correct datatypes

```{r}
data_total = as.data.frame(data_total)
num_col = c("temperature", "Longitude", "Latitude", "incidence")
int_col = c("dengue_cases","population_count","year")
date_col = c("time_column")

# checking for correct data-type
data_total[int_col] = lapply(data_total[int_col], as.integer)
data_total[num_col] = lapply(data_total[num_col], as.numeric)
data_total[date_col] = lapply(data_total[date_col], as.Date)

data_total = subset(data_total, select = c("data_names", "Reporting_areas", "year", "month", "time_column", "dengue_cases", "population_count", "temperature", "Longitude", "Latitude", "incidence")) # brings the columns into a better order

save(data_total, file = "data_total.RData")
summary(data_total)
```

## Creating a dataframe for mapping by setting values of Bungkan equal to Nong Khai

```{r}
map_data_total = data_total
for (k in 1:dim(data_total)[1]){
  Datum = data_total$time_column[k]
  
  if (data_total$Reporting_areas[k]=="Bungkan") {
    
    for(p in 1:dim(data_total)[1]){
      
      if(data_total$Reporting_areas[p] == "Nong Khai" & data_total$time_column[p] == Datum){
        map_data_total[k,c("temperature", "incidence", "dengue_cases")] = data_total[p,c("temperature", "incidence", "dengue_cases")]
      }
    }
  }
}
  
data_total = data_total[data_total$Reporting_areas != "Bungkan", ]
row.names(data_total) = NULL
```

## Creating a dataframe for monsoon season

```{r}
vec_month_monsoon = c("Jun", "Jul", "Aug", "Sep")
data_monsoon = data_total[data_total$month %in% vec_month_monsoon,]

rmv.rows.m = apply(data_monsoon,1,function(x){sum(is.na(x))})
i.missing.m = which(rmv.rows.m >0)
clean_data_monsoon = as.data.frame(data_monsoon[-i.missing.m,])
```

## Creating a dataframes for the total dengue cases in Thailand over each month

```{r}
month_vec = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec" ) 
total_thailand = data.frame(data_names = character(), dengue_cases = numeric())

for (i in 2006:2020) {
  name_years = paste0("rows_", i)
  assign(name_years, data_total[data_total$year == i, ]) 
  
  for (month in month_vec) {
    name_months = paste0(name_years, month)
    a = get(name_years)
    assign(name_months, a[a$month == month, ])
    
    b = get(name_months)
    total_cases = sum(b$dengue_cases, na.rm = T)
    name_column = paste0(i,"_", month)
    placeholder_df = data.frame(data_names = name_column, dengue_cases = total_cases)   
    total_thailand = rbind(total_thailand, placeholder_df)
    
  }
}

total_thailand = cbind(total_thailand, time_column_short)
```
