---
title: "Bioinfo_Projekt_Data_Description"
author: "Me"
date: "2023-06-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
summary(data_total)
```

# Temperature development in Thailand 2006 - 2022

Creating vector of mean temperature in Thailand over the years

```{r}
 mean_tempvec = c()
for (month in c(1:180)) {
   tempvec = c()
  for (district in c(1:76)) {
   temp = data_total$temperature[month + 180*(district-1)]
   tempvec = c(tempvec,temp)
  }
 mean_tempvec = c(mean_tempvec,mean(tempvec))
 }
mean_tempvec
```

**Plot of yearly mean temperature in Thailand over the years + linear regression of temperature**

```{r}

# Libraries
library(ggplot2)

# create data
month_dates = seq(as.Date("2006-01-01"), as.Date("2020-12-01"), by = "month", format = "%m")

xValue = month_dates
yValue = mean_tempvec
data = data.frame(xValue,yValue)

# Plot

ggplot() +
  geom_line(data, mapping=aes(x=xValue, y=yValue)) +
   stat_smooth(method = "lm", col = "red", mapping=aes(x=xValue, y=yValue)) +
  xlab("Years") +
  ylab("Temperature [°C]")
```

**Computing minimum and maximum average temperature**

```{r}
max_mean_tempvec = max(mean_tempvec)

min_mean_tempvec = min(mean_tempvec)

max_mean_tempvec
min_mean_tempvec


```

Function for plotting temperature in individual timeframe

```{r}

subset_mean_temp = function(start, end, mean_tempvec, month_dates){
  # input as string
  start_date = as.Date(start)
  end_date = as.Date(end)
  
  subset_temp = subset(mean_tempvec, month_dates >= start_date & month_dates <= end_date)
  
  month_dates_year = seq(start_date, end_date, by = "month", format = "%m")
  
  data_year = data.frame(month_dates_year, subset_temp)
  
  ggplot(data_year, mapping = aes(x=month_dates_year, y=subset_temp))+
    geom_line()+
    geom_point()+
    xlab("Months") +
    ylab("Temperature [°C]") +
    ggtitle(paste("Monthly mean temperature in Thailand", start, " - ", end))
}

```

Plot timeframe of min and max temperature

```{r}
subset_mean_temp("2013-06-01", "2014-06-30", mean_tempvec, month_dates)
subset_mean_temp("2016-01-01", "2016-12-31", mean_tempvec, month_dates)
```

# Correlation

```{r}
cor = cor(data_total$incidence, data_total$temperature, use = "pairwise.complete.obs")
plot(data_total$temperature, data_total$incidence, main = "incidence over temperature", xlab = "inzidenz", ylab = "temperature")
```

#ARIMA Dengue cases

```{r}
library(ggplot2)
library(forecast)

Ts_thai <- ts(total_thailand$dengue_cases, frequency = 12, start = c(2006, 1))
TSstudio::ts_plot(Ts_thai)

# moving average 
autoplot(Ts_thai) +
  autolayer(ma(Ts_thai, order = 12)) +
  xlab("Year") + ylab("dengue_cases") +
  ggtitle("Dengue_cases_thailand_with_MA") 

#additive decomposing
decomposed_thai_add <- decompose(Ts_thai, type="additive") # use type = "additive" for additive components
plot(decomposed_thai_add)
grid(lty = 2, col = "gray")
                      
hist(decomposed_thai_add$random, breaks = 20) 


#seasonal decomposing
decomposed_thai_seas <- stl(Ts_thai, s.window = "periodic")
plot(decomposed_thai_seas, main = "Decomposition of a periodic time series", xaxt = "n")
grid(lty = 2, col = "gray")


library(tseries)
adf.test(Ts_thai) # p-value (0.01) < 0.05 indicates the TS is stationary --> suggests it is stationary
# alternative hypothesis: stationary
kpss.test(Ts_thai, null = c("Level", "Trend"))

#both test indicate stationarity

acf_thailand <- acf(Ts_thai)

pacf_thailand <- pacf(Ts_thai)

```

##ARIMA differencing (not necessary as the time series is stationary) 

```{r}

Ts_thai_seasdiff <- diff(Ts_thai, lag = frequency(Ts_thai), differences = 1)  # seasonal differencing 
plot(Ts_thai_seasdiff, type="l", main = "Seasonally Differenced")  

Ts_thai_stat <- diff(Ts_thai_seasdiff, differences = 1)
plot(Ts_thai_stat, type = "l", main = "Differenced and Stationary")  

acfRes_stationary <- acf(Ts_thai_stat)
pacfRes_stationary <- pacf(Ts_thai_stat)
```

##ARIMA "by hand"

```{r}
p_range <- 0:4  # AR order
d_range <- 0:2 # I order
q_range <- 0:3 # MA order

# Initialize variables to store AIC and model information
best_aic <- Inf
best_model <- NULL
p_best <- 0
d_best <- 0
q_best <- 0

# Loop through the combinations of orders
for (p in p_range) {
  for (d in d_range) {
    for (q in q_range) {
      print(p); print(d); print(q)
      # Fit the ARIMA model
      model <- arima(Ts_thai, order = c(p, d, q)) 
      
      # Calculate AIC
      aic <- AIC(model)
      print(aic)
      # Check if the current model has lower AIC than the best model so far
      if (aic < best_aic) {
        best_aic <- aic
        best_model <- model
        p_best <- p
        d_best <- d
        q_best <- q
      }
    }
  }
}

# Print the best model and its AIC
best_model
best_aic; p_best; d_best; q_best

# Fit the best ARIMA model
best_model <- arima(Ts_thai, order = c(p_best, d_best, q_best))  # Replace with the best order values
best_model 

```

##Auto ARIMA

```{r}
library(forecast)
Ts_thai_auto  <- auto.arima(Ts_thai)
Ts_thai_auto

#auto_thailand_stat = auto.arima(Ts_thai_stat)
#auto_thailand_stat

#auto_thailand_2  <- auto.arima(Ts_thai, stepwise = F, approximation = F)
#auto_thailand_2
#this auto-ARIMA was removed because it took to long and had a higher AIC
```

```{r}
residuals_thai <- residuals(Ts_thai_auto)
plot(residuals_thai)

acf(residuals_thai)
hist(residuals_thai, breaks = 50)

portmanteau_thai <- Box.test(residuals_thai, lag = 12, type = "Ljung-Box")
portmanteau_thai 
# --> p value (0.41) > 0.05 --> no evidence for autocorrelation --> ARIMA model works well
```

```{r}
library(forecast)
horizon_thai <- 48 
forecasted_thai_auto <- forecast(Ts_thai_auto, h = horizon_thai)

forecasted_thai_Ts <- forecasted_thai_auto$mean
forecasted_thai_Ts  <- ts(forecasted_thai_Ts, frequency = 12, start = c(2021, 1) )
plot(forecasted_thai_Ts)
```

```{r}
forecast_thai_Df <- as.data.frame(forecasted_thai_auto)
forecast_thai_Df$time <- time(forecasted_thai_Ts)
names(forecast_thai_Df) <- c("point", "lo_80", "hi_80" , "lo_95", "hi_95", "time")
```

```{r}
library(ggplot2)
ggplot() +
  geom_line(data = Ts_thai, aes(x = time(Ts_thai), y = Ts_thai), color = "black") +
  geom_ribbon(data = forecast_thai_Df, aes(x = time, ymin = lo_95, ymax = hi_95), fill = "gray80") +
  geom_line(data = forecast_thai_Df, aes(x = time, y = point), color = "blue") +
  labs(title = "Time Series Forecast") +
  xlab("Year") +
  ylab("Dengue Cases") +
  theme_minimal()
```

##ARIMA forecast best model

```{r}
library(forecast)
forecasted_thai_best <- forecast(best_model, h = horizon_thai)

forecasted_Ts_thai_best <- forecasted_thai_best$mean
forecasted_Ts_thai_best  <- ts(forecasted_Ts_thai_best, frequency = 12, start = c(2021, 1) )
plot(forecasted_Ts_thai_best)
```

```{r}
forecast_thai_best_Df <- as.data.frame(forecasted_thai_best)
forecast_thai_best_Df$time <- time(forecasted_Ts_thai_best)
names(forecast_thai_best_Df) <- c("point", "lo_80", "hi_80" , "lo_95", "hi_95", "time")
```

```{r}
library(ggplot2)
ggplot() +
  geom_line(data = Ts_thai, aes(x = time(Ts_thai), y = Ts_thai), color = "black") +
  geom_ribbon(data = forecast_thai_best_Df, aes(x = time, ymin = lo_95, ymax = hi_95), fill = "gray80") +
  geom_line(data = forecast_thai_best_Df, aes(x = time, y = point), color = "blue") +
  labs(title = "Time Series Forecast") +
  xlab("Year") +
  ylab("Dengue Cases") +
  theme_minimal()
```

##ARIMA forecast 2016

```{r}
library(forecast)
Ts_thai_2016 = window(Ts_thai, end = c(2016, 12))

Ts_thai_auto_2016  <- auto.arima(Ts_thai_2016)
Ts_thai_auto_2016


forecasted_thai_2016 <- forecast(Ts_thai_auto_2016, h = horizon_thai)

forecasted_thai_2016_Ts <- forecasted_thai_2016$mean
forecasted_thai_2016_Ts  <- ts(forecasted_thai_2016_Ts, frequency = 12, start = c(2017, 1) )
plot(forecasted_thai_2016_Ts)

forecast_thai_2016_Df <- as.data.frame(forecasted_thai_2016)
forecast_thai_2016_Df$time <- time(forecasted_thai_2016_Ts)
names(forecast_thai_2016_Df) <- c("point", "lo_80", "hi_80" , "lo_95", "hi_95", "time")

library(ggplot2)
ggplot() +
  geom_ribbon(data = forecast_thai_2016_Df, aes(x = time, ymin = lo_95, ymax = hi_95), fill = "gray80") +
  geom_line(data = forecast_thai_2016_Df, aes(x = time, y = point), color = "blue") +
  geom_line(data = Ts_thai, aes(x = time(Ts_thai), y = Ts_thai), color = "black") +
  labs(title = "Time Series Forecast") +
  xlab("Year") +
  ylab("Dengue Cases") +
  theme_minimal()
```


##comparing the decomposition of temperature and cases

```{r}
library(ggplot2)
library(forecast)
library(gridExtra)

total_thailand_temp = cbind(total_thailand, mean_tempvec)
colnames(total_thailand_temp)[4] = "temperature"

Ts_thai_temp = ts(total_thailand_temp$temperature, frequency = 12, start = c(2006, 1))
TSstudio::ts_plot(Ts_thai_temp)

#additive decomposing of the temperature
decomposed_thai_temp_add = decompose(Ts_thai_temp, type="additive") 
plot(decomposed_thai_temp_add)
grid(lty = 2, col = "gray")

#additive decomposing of the dengue cases
decomposed_thai_add = decompose(Ts_thai, type="additive")
plot(decomposed_thai_add)
grid(lty = 2, col = "gray")

case_trend = decomposed_thai_add$trend
temp_trend = decomposed_thai_temp_add$trend

plot_case_trend = ggplot(case_trend, aes(x, y)) + geom_line() + labs(title = "Cases_trend")
plot_temp_trend = ggplot(temp_trend, aes(x, y)) + geom_line() + labs(title = "Temperature_trend")

plot_case_temp_trend = grid.arrange(plot_case_trend, plot_temp_trend, ncol = 1)
print(plot_case_temp_trend)


```


```{r}
case_seasonal = decomposed_thailand_add$seasonal
temp_seasonal = decomposed_thailand_add_temp$seasonal

plot_case_seasonal = ggplot(case_seasonal, aes(x, y)) + geom_line() + labs(title = "Cases_seasonal")
plot_temp_seasonal = ggplot(temp_seasonal, aes(x, y)) + geom_line() + labs(title = "Temperature_seasonal")

plot_case_temp_seasonal = grid.arrange(plot_case_seasonal, plot_temp_seasonal, ncol = 1)
print(plot_case_temp_trend)

```


# mean of dengue_cases and temperature of whole thailand per month

```{r}
library(ggplot2)
library(scales)

#dengue cases of whole thailand per month
f_mean_dengue = function(){
  mean_dengue = data.frame(dengue_cases = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(dengue_cases = mean(month$dengue_cases, na.rm = TRUE), time_column = month$time_column[1])
    mean_dengue = rbind(mean_dengue, m_month)
  }
  return(mean_dengue)
}


f_mean_temp = function(){
  mean_temp = data.frame(temperature = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(temperature = mean(month$temperature, na.rm = TRUE), time_column = month$time_column[1])
    mean_temp = rbind(mean_temp, m_month)
  }
  return(mean_temp)
}

plot1 = ggplot2::ggplot() +
  geom_line(data = f_mean_dengue(), aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases") +
  scale_x_continuous(breaks = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "2 year"))

plot2 = ggplot2::ggplot() +
  geom_line(data = f_mean_temp(), aes(x = time_column, y = temperature), color = "red") +
  labs(x = "time", y = "temperature") +
  scale_x_continuous(breaks = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "2 year"))


gridExtra::grid.arrange(plot1, plot2, nrow = 2)
```

## mean and median comparison

```{r}
library(ggplot2)

f_median_dengue = function(){
  median_dengue = data.frame(dengue_cases = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(dengue_cases = median(month$dengue_case, na.rm = TRUE), time_column = month$time_column[1])
    median_dengue = rbind(median_dengue, m_month)
  }
  return(median_dengue)
}

f_median_temp = function(){
  median_temp = data.frame(temperature = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(temperature = median(month$temperature, na.rm = TRUE), time_column = month$time_column[1])
    median_temp = rbind(median_temp, m_month)
  }
  return(median_temp)
}
#compute difference between mean and median
dif_mm_dengue = f_mean_dengue()["dengue_cases"]-f_median_dengue()["dengue_cases"]
dif_mm_dengue = cbind(dif_mm_dengue,f_mean_dengue()["time_column"])

plot_mean_dengue = ggplot2::ggplot() +
  geom_line(data = f_mean_dengue(), aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases", title = "Mean dengue cases")

plot_median_dengue = ggplot2::ggplot() +
  geom_line(data = f_median_dengue(), aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases", title = "Median dengue cases")

plot_dif_dengue = ggplot2::ggplot() +
  geom_line(data = dif_mm_dengue, aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases", title = "Difference between mean and median")


dif_mm_temp = f_mean_temp()["temperature"]-f_median_temp()["temperature"]
dif_mm_temp = cbind(dif_mm_temp,f_mean_temp()["time_column"])

plot_mean_temp = ggplot2::ggplot() +
  geom_line(data = f_mean_temp(), aes(x = time_column, y = temperature)) +
  labs(x = "time", y = "temperature", title = "Mean temperature")

plot_median_temp = ggplot2::ggplot() +
  geom_line(data = f_median_temp(), aes(x = time_column, y = temperature)) +
  labs(x = "time", y = "temperature", title = "Median temperature")

plot_dif_temp = ggplot2::ggplot() +
  geom_line(data = dif_mm_temp, aes(x = time_column, y = temperature)) +
  labs(x = "time", y = "temperature", title = "Difference between mean and median")

gridExtra::grid.arrange(plot_mean_dengue, plot_mean_temp, plot_median_dengue, plot_median_temp,plot_dif_dengue,plot_dif_temp, nrow = 3)

#wenn man nur die Temp-Plots haben möchte
#gridExtra::grid.arrange(plot_mean_temp, plot_median_temp, plot_dif_temp, nrow=3)
```

## mean of dengue_cases and temperature of whole thailand per year

```{r}
f_mean_dengue.y.t = function(){
  mean_dengue.y.t = data.frame(dengue_cases = numeric(), time_column = as.Date(character()))
  for (i in c(2006:2020)) {
    month = data_total[data_total$year == i,]
    m_month = data.frame(dengue_cases = mean(month$dengue_cases, na.rm = TRUE), time_column = month$time_column[1])
    mean_dengue.y.t = rbind(mean_dengue.y.t, m_month)
  }
  return(mean_dengue.y.t)
}
mean_dengue.y.t = as.data.frame(f_mean_dengue.y.t())

f_mean_temp.y.t = function(){
  mean_temp.y.t = data.frame(temperature = numeric(), time_column = as.Date(character()))
  for (i in c(2006:2020)) {
    month = data_total[data_total$year == i,]
    m_month = data.frame(temperature = mean(month$temperature, na.rm = TRUE), time_column = month$time_column[1])
    mean_temp.y.t = rbind(mean_temp.y.t, m_month)
  }
  return(mean_temp.y.t)
}
mean_temp.y.t = as.data.frame(f_mean_temp.y.t())

```

#Plot für Bangkok

```{r}
dengue_plot_cases = c()
dengue_plot_time = c()

dengue_plot_cases = data_total$dengue_cases[data_total$Reporting_areas == "Bangkok"]
dengue_plot_time = data_total$time_column[data_total$Reporting_areas == "Bangkok"]

plot(dengue_plot_time,dengue_plot_cases, type = "o")
```

#distribution of incidence and temperature

```{r}

q.dengue = quantile(data_total$incidence, probs = c(0.01,0.1,0.25,0.5,0.75,0.9,0.99), na.rm = TRUE)
hist(data_total$incidence,  breaks = 200, main = "Distribution of incidence", xlab = "incidence", freq = F);abline(v=q.dengue,lty=3,lwd=2,col='red')
lines(density(data_total$incidence, na.rm = T), col = "blue", lwd = 2)

q.temp = quantile(data_total$temperature, probs = c(0.01,0.1,0.25,0.5,0.75,0.9,0.99), na.rm = TRUE)
hist(data_total$temperature,  breaks = 200, main = "Distribution of temperature", xlab = "temperature", freq = F);abline(v=q.temp,lty=3,lwd=2,col='red')
curve(dnorm(x, mean = mean(data_total$temperature, na.rm = TRUE), sd = sd(data_total$temperature, na.rm = TRUE)), 
      add = TRUE, col = "green", lwd = 2)
lines(density(data_total$temperature, na.rm = T), col = "blue", lwd = 2)

```

#correlation between temperature and incidence
```{r}
library(npreg)
library(ggplot2)

smoothed_curve <- ss(clean_data_monsoon_5.95$temperature, clean_data_monsoon_5.95$incidence, nknots = 10)

plot(clean_data_monsoon$temperature, clean_data_monsoon$incidence, pch = 20,
     ylab = "incidence",
     xlab = "temperature",
  abline(v = mean(clean_data_monsoon$temperature), col = "red"))
plot(clean_data_total$temperature, clean_data_total$incidence, pch = 20,
     ylab = "incidence",
     xlab = "temperature",
  abline(v = mean(clean_data_total$temperature), col = "red"))

plot(smoothed_curve, col = "blue")


plot(rank(data_total$temperature), rank(data_total$incidence), pch = 20,
     ylab = "incidence",
     xlab = "temperature",
  abline(v = mean(data_total$temperature), col = "red"))

qqplot(data_total$temperature, data_total$incidence)

#pearson correlation y.t.
cor(mean_dengue.y.t$dengue_cases, mean_temp.y.t$temperature)
#spearman correlation y.t.
cor(mean_dengue.y.t$dengue_cases, mean_temp.y.t$temperature, method = "spearman")

#pearson correlation m.p.
cor(clean_data_total$dengue_cases, clean_data_total$temperature)
#spearman correlation m.p.
cor(clean_data_total$dengue_cases, clean_data_total$temperature, method = "spearman")


pheatmap::pheatmap(cor(clean_data_total[c("dengue_cases", "population_count", "temperature", "incidence")]))

```
#percentage of dengue cases between 27 and 29,5 celsius
```{r}
data_27_29.5 = data_total[data_total$temperature >= 27 & data_total$temperature <= 29.5,]
sum(data_27_29.5$dengue_cases, na.rm = T)/sum(data_total$dengue_cases, na.rm = T)


```

## correlation between different areas

```{r}
#mean of temp and incidence over all years per area
t_data_total = t(data_total)
r_areas = unique(data_total$Reporting_areas)
m_inc_temp.y.r = data.frame(Reporting_areas = character(), incidence = numeric(), temperature = numeric())
for (i in r_areas){
  temp_inc = apply(data_total[which(clean_data_total$Reporting_areas == i),c("incidence", "temperature")], 2,function(g){mean(g, na.rm = TRUE)})
  temp_inc_df = data.frame(Reporting_areas = i, incidence = temp_inc["incidence"], temperature = temp_inc["temperature"])
  m_inc_temp.y.r = rbind(m_inc_temp.y.r, temp_inc_df)
  rownames(m_inc_temp.y.r) = 1:nrow(m_inc_temp.y.r)
}

```

# k-means clustering

```{r}
#elbow method
km = kmeans(m_inc_temp.y.r["incidence"], centers = 2, nstart = 10)
km$tot.withinss

wss = sapply(2:7,function(k) { 
  kmeans(m_inc_temp.y.r["incidence"], centers = k)$tot.withinss
})
plot(2:7,wss,type='b',pch=19,xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")


#silhoutte method
library(cluster)
#library (NbClust)
#library (clustertend)
library (factoextra)
#library (fpc)
#library (clValid)

#find optimal number of clusters by incidence
D_inc = dist(m_inc_temp.y.r["incidence"])
km_inc = kmeans(m_inc_temp.y.r["incidence"], centers = 4, nstart = 10)
s_inc = silhouette(km_inc$cluster,D_inc)
plot(s_inc)
fviz_nbclust(m_inc_temp.y.r["incidence"], pam, method = "silhouette")+ theme_classic()

#find optimal number of clusters by temperature
D_temp = dist(m_inc_temp.y.r["temperature"])
km_temp = kmeans(m_inc_temp.y.r["temperature"], centers = 4, nstart = 10)
s_temp = silhouette(km_temp$cluster,D_temp)
plot(s_temp)
fviz_nbclust(m_inc_temp.y.r["temperature"], pam, method = "silhouette")+ theme_classic()

#find optimal number of clusters by temperature and incidence
D_inc_temp = dist(scale(m_inc_temp.y.r[,c("incidence","temperature")]))
km_inc_temp = kmeans(scale(m_inc_temp.y.r[,c("incidence","temperature")]), centers = 3)
s_inc_temp = silhouette(km_inc_temp$cluster,D_inc_temp)
plot(s_inc_temp)
fviz_nbclust(scale(m_inc_temp.y.r[,c("incidence","temperature")]), pam, method = "silhouette")+ theme_classic()

m_inc_temp_cluster.y.r = data.frame(m_inc_temp.y.r, km_inc_temp$cluster)

plot1 = ggplot(m_inc_temp.y.r, aes(x = temperature, y = incidence, color = as.factor(km_inc$cluster))) +
  geom_point() +
  labs(title = "Partitioning Clustering Plot by incidence") +
  theme_classic()

plot2 = ggplot(m_inc_temp.y.r, aes(x = temperature, y = incidence, color = as.factor(km_temp$cluster))) +
  geom_point() +
  labs(title = "Partitioning Clustering Plot by temperature") +
  theme_classic()

plot3 = ggplot(m_inc_temp.y.r, aes(x = temperature, y = incidence, color = as.factor(km_inc_temp$cluster))) +
  geom_point() +
  labs(title = "Partitioning Clustering Plot") +
  theme_classic()
plot3
gridExtra::grid.arrange(plot1, plot2, plot3, ncol = 1)
```

## probieren mit time series Objekten

```{r}
my_ts = ts(data_total$dengue_cases, start = c(data_total$time_column[1]), frequency = 12)
head(my_ts)
TSstudio::ts_plot(my_ts)
```

#GAM
```{r}
library(readxl)
library(nlme)
library(gam)
library(sdm)
library(raster)
library(ncdf4)
library(mgcv)

attach(data_monsoon)
model1 = gam(incidence~s(temperature, k = 5), family = "quasipoisson", na.rm = T)
plot.gam(model1)
summary(model1) #edf: → a value close to 1 tends to be close to a linear term → a higher value means that the spline is more wiggly (non-linear). 
#F-Wert und der p-Wert geben die statistische Signifikanz der glatten Funktion an. In diesem Fall ist die glatte Funktion von "temperature" hoch signifikant (p-Wert < 2e-16), was darauf hindeutet, dass sie einen signifikanten Einfluss auf die Vorhersage der "incidence" hat.Der angegebene R-squared-Wert (R-sq.(adj)) von 0.0407 gibt an, wie gut das Modell die Varianz der abhängigen Variable erklärt. In diesem Fall erklärt das Modell etwa 4,07% der Varianz in den Daten. Ein niedriger GCV-Wert deutet auf eine gute Anpassung des Modells hin.
k.check(model1)
gam.check(model1)

AIC_best = c()
for (i in 3:10){
  smooth_model = gam(incidence~s(temperature, k = i))
  linear_model = gam(incidence~temperature, k = i)
  AIC_best = c(AIC_best, AIC(linear_model, smooth_model))
}
AIC_best
detach(data_monsoon)
```

## GAM forecast

```{r, fig.width = 10, fig.height = 15}
library(raster)
library(gam)
library(sdm)
library(readxl)
library(mgcv)
library(ncdf4)

## lower_limit = quantile(data_monsoon$temperature, probs = 0.05, na.rm = T)
#upper_limit = quantile(data_monsoon$temperature, probs = 0.95, na.rm = T)

#data_monsoon_5.95 = subset(data_monsoon, temperature >= lower_limit & temperature <= upper_limit)

attach(data_total)

model2 <- gam(incidence~s(temperature, k = 16), family = "quasipoisson")
tas <- stack("tas_SEA22_MPI_rcp85_2021-2040_grid_subc_daymean_monmean_timmean.nc")
#tas_test = crop(tas, my.area)
names(tas) <- "temperature"
p1 <- raster::predict(tas, model2, type = "response")

thailand <- raster::getData('GADM', country='THA', level=1)
my.area <- extent(thailand)
my.area = extent(95,107,5.5,20.5)
my.p1 <- crop(p1, my.area)
my.p1.mask <- mask(my.p1, thailand)


plot(my.p1.mask, legend.width=2, legend.shrink=0.5, axes= F, box=F, zlim = c(0,22.8)) + plot(thailand, add=TRUE)


detach(data_total)
```

#average incidence 2006 to 2020 map
```{r}
library(ggplot2)
library(sf)

#function to plot the incidence of a specific period and over a specific area
f_map_mean_inc.y.p = function(period, areas){
  map_mean_inc.y.p <- data.frame(incidence = numeric(), time_column = as.Date(character()), Reporting_areas = character())
  for (i in period){
    year = map_data_total[map_data_total$year == i,]
    for (p in areas) {
      provinz = year[year$Reporting_areas == p,]
      m_y_p = data.frame(incidence = mean(provinz$incidence, na.rm = TRUE), time_column = provinz$time_column[1], Reporting_areas = provinz$Reporting_areas[1])
      map_mean_inc.y.p = rbind(map_mean_inc.y.p, m_y_p)
    }
  }
  return(map_mean_inc.y.p)
}

#function to plot the incidence the period from 2006 to 2020 for specific areas
f_map_mean_inc.p = function(areas){
  map_mean_inc.p <- data.frame(time_column = as.Date(character()), Reporting_areas = character(), incidence = numeric())
  for (p in areas) {
      provinz = map_data_total[map_data_total$Reporting_areas == p,]
      m_p = data.frame(time_column = provinz$time_column[1], Reporting_areas = provinz$Reporting_areas[1],incidence = mean(provinz$incidence, na.rm = TRUE))
      map_mean_inc.p = rbind(map_mean_inc.p, m_p)
  }
  return(map_mean_inc.p)
}


r_areas2 = unique(data_total$Reporting_areas)
mean_inc.p <- data.frame(time_column = as.Date(character()), Reporting_areas = character(), incidence = numeric())
  for (p in r_areas2) {
      provinz = data_total[data_total$Reporting_areas == p,]
      m_p = data.frame(time_column = provinz$time_column[1], Reporting_areas = provinz$Reporting_areas[1],incidence = mean(provinz$incidence, na.rm = TRUE))
      mean_inc.p = rbind(mean_inc.p, m_p)
}

map_mean_inc.p = f_map_mean_inc.p(areas = unique(map_data_total$Reporting_areas))
district_sf = st_read("gadm36_THA_shp/gadm36_THA_1.shp")
district_sf$incidence <- map_mean_inc.p$incidence


ggplot2::ggplot() +
  geom_sf(data = district_sf, aes(fill = incidence)) +
  scale_fill_gradientn(colors = rev(terrain.colors(100)), limits = c(0,22.8)) + labs(fill = "Incidence") + labs(title = "Average incidences for 2006-2020 ")

```

Funktion für Inzidenz plot der Monate

```{r}

library(sf)
library(ggplot2)


district_sf = st_read("gadm36_THA_shp/gadm36_THA_1.shp")


map_incidence_monthly = function(incidence_month){
  
 max_incidence = which.max(incidence_month$incidence)
max_incidence_area = incidence_month$Reporting_areas[max_incidence]



ggplot() +
  geom_sf(data = district_sf, aes(fill = incidence_month$incidence)) +
  scale_fill_gradient(low = "yellow", high = "red") +
  labs(fill = paste("Highest incidence =", max_incidence, "in", max_incidence_area))
  
  
}
#input for incidence month in format: rows_YYYY_Mon



map_incidence_monthly(rows_2016Apr)


```
