---
title: "Bioinfo_Projekt_Einlesen"
author: "Frederik Racky"
date: "2023-05-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("readxl")
library(readxl)
```

##Einlesen der Tabellen und umwandeln der Tabellen in data frames

```{r}
library(readxl)
library(dplyr)

Endung <- "_dengue_extracted.xlsx"
Anfang <- "Dengue"

for (i in 2006:2020) {
  
  Dateiname <- paste0(i, Endung) #erstellen das Dateinamens
  Name_df <- paste0(Anfang, i) 
  assign(Name_df, read_excel(Dateiname)) #erstellen der Variable und zuweisen der Excel-Tabelle
  
  df_dengue <- get(Name_df)
  vec_data_names <- vector("character")
  
  vec_data_names <- sapply(df_dengue$Reporting_areas, function(g) {
    paste(g[1], i)
  })
  data_names <- matrix(vec_data_names,nrow = 77, ncol = 1)
  df_dengue <- cbind(data_names, df_dengue)
  assign(Name_df,df_dengue)
}
total_years = data.frame()
for (i in 2006:2020){
  a = get(paste0("Dengue",i))
  total_years = rbind(total_years, a)
}
```

#Rows als Zeitachse mit Monaten

```{r}
m = 1
data_total = data.frame(col1 = character(0), col2 = character(0), col3 = numeric(0), month = character(0))
colnames(data_total) = c("data_names", "Reporting_areas", "dengue_cases", "month")
sorted_total_years = total_years[order(total_years$"Reporting_areas"),]
for (k in 1:1155){
  for (j in 3:14){
    square <- sorted_total_years[k, c(1, 2, j)]
    month = colnames(square[3])
    square <- cbind(square,month)
    colnames(square) = c("data_names", "Reporting_areas", "dengue_cases", "month")
    data_total <- rbind(data_total, square)
    colnames(data_total) = c("data_names", "Reporting_areas", "dengue_cases", "month")
    data_total[m,1] = paste(data_total[m,1], month[1])
    m = m + 1
  }
}
```

#adding temperature column

```{r}
m = 1
temp_data = read_excel("era5_data_2006_2020_thailand_monmean.xlsx")
temp_data = as.data.frame(temp_data)
temp_data_sorted = data.frame( col1 = numeric(0), Longitude = numeric(0), Latitude = numeric(0))
for (i in 1:77) {
  for (j in 5:184) {
    temp_data_sorted[m,1] = temp_data[i,j]
    temp_data_sorted[m,2] = temp_data[i,3]
    temp_data_sorted[m,3] = temp_data[i,4]
    m = m + 1
  }
}
data_total = cbind(data_total, temp_data_sorted) 
colnames(data_total)[5] = "temperature"
```

#Einlesen Population data

```{r}
library(readxl)

Endung <- "_population.xlsx"
#empty dataframe for the data
total_population = data.frame(Reporting_areas = character(), population_count = numeric())

# stack years of 2006-2011 below each other
for (i in 2006:2011) {
  population_df  = data.frame()
  Dateiname <- paste0(i, Endung) #erstellen das Dateinamens
  population_df = read_excel(Dateiname) #erstellen der Variable und zuweisen der Excel-Tabelle
  vec_years = rep(i, times = 77)
  population_df = cbind(population_df,vec_years)
  names(population_df) = c("Reporting_areas", "population_count", "year")
  # multiply every row 12 times for each month
  repeat_count <- 12
  population_df_replicated <- population_df[rep(row.names(population_df), each = repeat_count), ]
  row.names(population_df_replicated) <- NULL
  
  #put all the years together
  total_population = rbind(total_population, population_df_replicated)
  
}

# stack years of 2012-2020 below each other and add to data_total
population_df = read_excel("2012-2020_population.xlsx")
colnames(population_df) = c("Reporting_areas", "population_count", "population_count", "population_count", "population_count", "population_count","population_count","population_count","population_count","population_count")
pop_placeholder = data.frame(Reporting_areas = character(), population_count = numeric(), year = numeric())
col_num = 2
for (i in 2012:2020) {
  year = rep(i, times = 77)
  pop_placeholder = rbind(pop_placeholder, cbind(population_df[,c(1,col_num)],year))
  names(pop_placeholder) = c("Reporting_areas", "population_count", "year")
  col_num = col_num + 1
}

# multiply every row 12 times for each month
repeat_count <- 12
population_df_replicated <- pop_placeholder[rep(row.names(pop_placeholder), each = repeat_count), ]
row.names(population_df_replicated) <- NULL

# combine all population data in each 
total_population = rbind(total_population, population_df_replicated)
sorted_total_population = total_population[order(total_population$"Reporting_areas"),]

data_total = cbind(data_total,sorted_total_population[,c(2,3)])
row.names(data_total) = NULL
```

# adding time column

```{r}
first_date = as.Date("2006-01-01")
last_date = as.Date("2020-12-31")

time_column_short = seq(first_date, last_date, by = "month")
time_column_short = as.Date(time_column_short, format = "%Y-%m")
time_column = rep(time_column_short, times = 77)

data_total = cbind(data_total, time_column)
#data_total = subset(data_total, select = -time_column) 
```

#Detect missing values

```{r}
detect_na <- function(x) {
  is_na <- x %in% c("N/A", "na", "NA", "n/a")
  replace(x, is_na, NA)
}
data_total <- apply(data_total, 2, detect_na)
rmv.rows = apply(data_total,1,function(x){sum(is.na(x))})
```

#struture data frame and convert into right datatypes

```{r}
data_total = as.data.frame(data_total)
num_col = c("temperature", "Longitude", "Latitude")
int_col = c("dengue_cases","population_count","year")
data_total[int_col] = lapply(data_total[int_col], as.integer)
data_total[num_col] = lapply(data_total[num_col], as.numeric)
data_total = subset(data_total, select = c("data_names", "Reporting_areas", "year", "month", "time_column", "dengue_cases", "population_count", "temperature", "Longitude", "Latitude"))
```

# creating a dataframe for the total dengue cases in thailand over the months

```{r}
month_vec = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec" ) 
total_thailand = data.frame(data_names = character(), dengue_cases = numeric())

for (i in 2006:2020) {
  name_years = paste0("rows_", i)
  assign(name_years, data_total[data_total$year == i, ]) 
  
  for (month in month_vec) {
    name_months = paste0(name_years, month)
    a = get(name_years)
    assign(name_months, a[a$month == month, ])
    
    b = get(name_months)
    total_cases = sum(b$dengue_cases, na.rm = T)
    name_column = paste0(i,"_", month)
    placeholder_df = data.frame(data_names = name_column, dengue_cases = total_cases)   
    total_thailand = rbind(total_thailand, placeholder_df)
    
  }
}

total_thailand = cbind(total_thailand, time_column_short)
total_thailand$dengue_cases = as.numeric(total_thailand$dengue_cases)
```

# moving average 
autoplot(Ts_thailand) +
  autolayer(ma(Ts_thailand, order = 12)) +
  xlab("Year") + ylab("dengue_cases") +
  ggtitle("Dengue_cases_thailand_with_MA") 


#ARIMA Dengue cases

```{r}
library(ggplot2)
library(forecast)

Ts_thai <- ts(total_thailand$dengue_cases, frequency = 12, start = c(2006, 1))
TSstudio::ts_plot(Ts_thai)

# moving average 
autoplot(Ts_thai) +
  autolayer(ma(Ts_thai, order = 12)) +
  xlab("Year") + ylab("dengue_cases") +
  ggtitle("Dengue_cases_thailand_with_MA") 

#additive decomposing
decomposed_thai_add <- decompose(Ts_thai, type="additive") # use type = "additive" for additive components
plot(decomposed_thai_add)
grid(lty = 2, col = "gray")
                      
#seasonal decomposing
decomposed_thai_seas <- stl(Ts_thai, s.window = "periodic")
plot(decomposed_thai_seas, main = "Decomposition of a periodic time series", xaxt = "n")
grid(lty = 2, col = "gray")


library(tseries)
adf.test(Ts_thai) # p-value (0.01) < 0.05 indicates the TS is stationary --> suggests it is stationary
# alternative hypothesis: stationary
kpss.test(Ts_thai, null = c("Level", "Trend"))

#both test indicate stationarity

acf_thailand <- acf(total_thailand$dengue_cases)

pacf_thailand <- pacf(total_thailand$dengue_cases)

```

#ARIMA differencing
```{r}
# 1st Seasonal Differencing 
# purpose of seasonal differencing is to stabilize the mean and variance of the ts
Ts_thai_seasdiff <- diff(Ts_thai, lag = frequency(Ts_thai), differences = 1)  # seasonal differencing 
plot(Ts_thai_seasdiff, type="l", main = "Seasonally Differenced")  

# 2md Seasonal Differencing
Ts_thai_stat <- diff(Ts_thai_seasdiff, differences = 1)
plot(Ts_thai_stat, type = "l", main = "Differenced and Stationary")  # --> USUALLY you do not difference more than TWO times 

acfRes_stationary <- acf(Ts_thai_stat)
pacfRes_stationary <- pacf(Ts_thai_stat)
```

#ARIMA "by hand"

```{r}
p_range <- 0:4  # AR order
d_range <- 0:2 # I order
q_range <- 0:3 # MA order

# Initialize variables to store AIC and model information
best_aic <- Inf
best_model <- NULL
p_best <- 0
d_best <- 0
q_best <- 0

# Loop through the combinations of orders
for (p in p_range) {
  for (d in d_range) {
    for (q in q_range) {
      print(p); print(d); print(q)
      # Fit the ARIMA model
      model <- arima(Ts_thai, order = c(p, d, q)) 
      
      # Calculate AIC
      aic <- AIC(model)
      print(aic)
      # Check if the current model has lower AIC than the best model so far
      if (aic < best_aic) {
        best_aic <- aic
        best_model <- model
        p_best <- p
        d_best <- d
        q_best <- q
      }
    }
  }
}

# Print the best model and its AIC
best_model
best_aic; p_best; d_best; q_best

# Fit the best ARIMA model
best_model <- arima(Ts_thai, order = c(p_best, d_best, q_best))  # Replace with the best order values
best_model 

```
#Auto ARIMA

```{r}
library(forecast)
Ts_thai_auto  <- auto.arima(Ts_thai)
Ts_thai_auto

#auto_thailand_stat = auto.arima(Ts_thai_stat)
#auto_thailand_stat

#auto_thailand_2  <- auto.arima(Ts_thai, stepwise = F, approximation = F)
#auto_thailand_2
#this auto-ARIMA was removed because it took to long and had a higher AIC
```


```{r}
residuals_thai <- residuals(Ts_thai_auto)
plot(residuals_thai)

acf(residuals_thai)

portmanteau_thai <- Box.test(residuals_thai, lag = 12, type = "Ljung-Box")
portmanteau_thai 
# --> p value (0.52) > 0.05 --> no evidence for autocorrelation --> ARIMA model works well
```

```{r}
library(forecast)
horizon_thai <- 48 
forecasted_thai_auto <- forecast(Ts_thai_auto, h = horizon_thai)

forecasted_thai_Ts <- forecasted_thai_auto$mean
forecasted_thai_Ts  <- ts(forecasted_thai_Ts, frequency = 12, start = c(2021, 1) )
plot(forecasted_thai_Ts)
```

```{r}
forecast_thai_Df <- as.data.frame(forecasted_thai_auto)
forecast_thai_Df$time <- time(forecasted_thai_Ts)
names(forecast_thai_Df) <- c("point", "lo_80", "hi_80" , "lo_95", "hi_95", "time")
```

```{r}
library(ggplot2)
ggplot() +
  geom_line(data = Ts_thai, aes(x = time(Ts_thai), y = Ts_thai), color = "black") +
  geom_ribbon(data = forecast_thai_Df, aes(x = time, ymin = lo_95, ymax = hi_95), fill = "gray80") +
  geom_line(data = forecast_thai_Df, aes(x = time, y = point), color = "blue") +
  labs(title = "Time Series Forecast") +
  xlab("Year") +
  ylab("Dengue Cases") +
  theme_minimal()
```

#ARIMA forecast best model

```{r}
library(forecast)
forecasted_thai_best <- forecast(best_model, h = horizon_thai)

forecasted_Ts_thai_best <- forecasted_thai_best$mean
forecasted_Ts_thai_best  <- ts(forecasted_Ts_thai_best, frequency = 12, start = c(2021, 1) )
plot(forecasted_Ts_thai_best)
```

```{r}
forecast_thai_best_Df <- as.data.frame(forecasted_thai_best)
forecast_thai_best_Df$time <- time(forecasted_Ts_thai_best)
names(forecast_thai_best_Df) <- c("point", "lo_80", "hi_80" , "lo_95", "hi_95", "time")
```

```{r}
library(ggplot2)
ggplot() +
  geom_line(data = Ts_thai, aes(x = time(Ts_thai), y = Ts_thai), color = "black") +
  geom_ribbon(data = forecast_thai_best_Df, aes(x = time, ymin = lo_95, ymax = hi_95), fill = "gray80") +
  geom_line(data = forecast_thai_best_Df, aes(x = time, y = point), color = "blue") +
  labs(title = "Time Series Forecast") +
  xlab("Year") +
  ylab("Dengue Cases") +
  theme_minimal()
```

#ARIMA forecast 2018
```{r}
library(forecast)
Ts_thai_2018 = window(Ts_thai, end = c(2018, 12))

Ts_thai_auto_2018  <- auto.arima(Ts_thai_2018)
Ts_thai_auto_2018


forecasted_thai_2018 <- forecast(Ts_thai_auto_2018, h = horizon_thai)

forecasted_thai_2018_Ts <- forecasted_thai_2018$mean
forecasted_thai_2018_Ts  <- ts(forecasted_thai_2018_Ts, frequency = 12, start = c(2018, 1) )
plot(forecasted_thai_2018_Ts)

forecast_thai_2018_Df <- as.data.frame(forecasted_thai_2018)
forecast_thai_2018_Df$time <- time(forecasted_thai_2018_Ts)
names(forecast_thai_2018_Df) <- c("point", "lo_80", "hi_80" , "lo_95", "hi_95", "time")

library(ggplot2)
ggplot() +
  geom_ribbon(data = forecast_thai_2018_Df, aes(x = time, ymin = lo_95, ymax = hi_95), fill = "gray80") +
  geom_line(data = forecast_thai_2018_Df, aes(x = time, y = point), color = "blue") +
  geom_line(data = Ts_thai, aes(x = time(Ts_thai), y = Ts_thai), color = "black") +
  labs(title = "Time Series Forecast") +
  xlab("Year") +
  ylab("Dengue Cases") +
  theme_minimal()
```


#comparing the decomposition of temperature and cases

```{r}
library(ggplot2)
library(forecast)
library(gridExtra)

total_thailand_temp = cbind(total_thailand, mean_tempvec)
colnames(total_thailand_temp)[4] = "temperature"

Ts_thai_temp <- ts(total_thailand_temp$temperature, frequency = 12, start = c(2006, 1))
TSstudio::ts_plot(Ts_thai_temp)

#additive decomposing of the temperature
decomposed_thai_temp_add <- decompose(Ts_thai_temp, type="additive") 
plot(decomposed_thai_temp_add)
grid(lty = 2, col = "gray")

#additive decomposing of the dengue cases
decomposed_thai_add <- decompose(Ts_thai, type="additive")
plot(decomposed_thai_add)
grid(lty = 2, col = "gray")

case_trend <- decomposed_thai_add$trend
temp_trend <- decomposed_thai_temp_add$trend

plot_case_trend <- ggplot(case_trend, aes(x, y)) + geom_line() + labs(title = "Cases_trend")
plot_temp_trend <- ggplot(temp_trend, aes(x, y)) + geom_line() + labs(title = "Temperature_trend")

plot_case_temp_trend <- grid.arrange(plot_case_trend, plot_temp_trend, ncol = 1)
print(plot_case_temp_trend)


```


```{r}
case_seasonal <- decomposed_thailand_add$seasonal
temp_seasonal <- decomposed_thailand_add_temp$seasonal

plot_case_seasonal <- ggplot(case_seasonal, aes(x, y)) + geom_line() + labs(title = "Cases_seasonal")
plot_temp_seasonal <- ggplot(temp_seasonal, aes(x, y)) + geom_line() + labs(title = "Temperature_seasonal")

plot_case_temp_seasonal <- grid.arrange(plot_case_seasonal, plot_temp_seasonal, ncol = 1)
print(plot_case_temp_trend)

```

#keine ahnung

```{r}
dimensions= dim(data_total)
R= dimensions[1]

for(Zeile in 1:R){
  Population <- data_total$population_count[Zeile]
  Infection <- data_total$dengue_cases[Zeile]
  Incidence = (Infection/Population)*100000
  data_total$incidence[Zeile] <- Incidence
}
```

# Temperature development in Thailand 2006 - 2022

Creating vector of mean temperature in Thailand over the years

```{r}
 mean_tempvec <- c()
for (month in c(1:180)) {
   tempvec <- c()
  for (district in c(1:77)) {
   temp <- data_total$temperature[month + 180*(district-1)]
   tempvec <- c(tempvec,temp)
  }
 mean_tempvec <- c(mean_tempvec,mean(tempvec))
 }
mean_tempvec
```

Plot of yearly mean temperature in Thailand over the years + linear regression of temperature

```{r}

# Libraries
library(ggplot2)

# create data
xValue <- month_dates
yValue <- mean_tempvec
data <- data.frame(xValue,yValue)

# Plot

ggplot() +
  geom_line(data, mapping=aes(x=xValue, y=yValue)) +
   stat_smooth(method = "lm", col = "red", mapping=aes(x=xValue, y=yValue)) +
  xlab("Years") +
  ylab("Temperature [°C]")
```

# Correlation

```{r}
cor = cor(data_total$incidence, data_total$temperature, use = "pairwise.complete.obs")
plot(data_total$temperature, data_total$incidence, main = "incidence over temperature", xlab = "inzidenz", ylab = "temperature")
```

#mean of dengue_cases of whole thailand

```{r}
library(ggplot2)
library(scales)


f_mean_dengue = function(){
  mean_dengue <- data.frame(dengue_cases = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(dengue_cases = mean(month$dengue_case, na.rm = TRUE), time_column = month$time_column[1])
    mean_dengue = rbind(mean_dengue, m_month)
  }
  return(mean_dengue)
}
  
f_mean_temp = function(){
  mean_temp <- data.frame(temperature = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(temperature = mean(month$temperature, na.rm = TRUE), time_column = month$time_column[1])
    mean_temp = rbind(mean_temp, m_month)
  }
  return(mean_temp)
}

plot1 = ggplot2::ggplot() +
  geom_line(data = f_mean_dengue(), aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases") +
  scale_x_date(breaks = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "2 year"))

plot2 = ggplot2::ggplot() +
  geom_line(data = f_mean_temp(), aes(x = time_column, y = temperature), color = "red") +
  labs(x = "time", y = "temperature") +
  scale_x_date(breaks = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "2 year"))

gridExtra::grid.arrange(plot1, plot2, nrow = 2)
```

#mean and median comparison

```{r}
f_median_dengue = function(){
  median_dengue <- data.frame(dengue_cases = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(dengue_cases = median(month$dengue_case, na.rm = TRUE), time_column = month$time_column[1])
    median_dengue = rbind(median_dengue, m_month)
  }
  return(median_dengue)
}

f_median_temp = function(){
  median_temp <- data.frame(temperature = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(temperature = median(month$temperature, na.rm = TRUE), time_column = month$time_column[1])
    median_temp = rbind(median_temp, m_month)
  }
  return(median_temp)
}


plot_mean_temp = ggplot2::ggplot() +
  geom_line(data = f_mean_temp(), aes(x = time_column, y = temperature)) +
  labs(x = "time", y = "temperature", title = "Mean temperature")

plot_mean_dengue = ggplot2::ggplot() +
  geom_line(data = f_mean_dengue(), aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases", title = "Mean dengue cases")

plot_median_temp = ggplot2::ggplot() +
  geom_line(data = f_median_temp(), aes(x = time_column, y = temperature)) +
  labs(x = "time", y = "temperature", title = "Median temperature")

plot_median_dengue = ggplot2::ggplot() +
  geom_line(data = f_median_dengue(), aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases", title = "Median dengue cases")

gridExtra::grid.arrange(plot_mean_dengue, plot_mean_temp, plot_median_dengue, plot_median_temp, nrow = 2)
```

#Plot für Bangkok

```{r}
dengue_plot_cases = c()
dengue_plot_time = c()

dengue_plot_cases <- data_total$dengue_cases[data_total$Reporting_areas == "Bangkok"]
dengue_plot_time <- data_total$time_column[data_total$Reporting_areas == "Bangkok"]

plot(dengue_plot_time,dengue_plot_cases, type = "o")
```

#correlation between temperature and dengue cases

```{r}
i.missing = which(rmv.rows > 0)


plot(data_total$temperature, data_total$incidence, pch = 20,
     ylab = "incidence",
     xlab = "mean temperature",
     lines(data_total$temperature, data_total$incidence, col = "red")
     )

```

#probieren mit time series Objekten

```{r}
my_ts <- ts(data_total$dengue_cases, start = c(data_total$time_column[1]), frequency = 12)
head(my_ts)
TSstudio::ts_plot(my_ts)
```

#GAM

```{r}

library(readxl)
library(nlme)
library(gam)
library(sdm)
library(raster)
library(ncdf4)
library(mgcv)

attach(data_total)
model1 <- gam(dengue_cases~s(temperature, k=6), family="quasipoisson")
plot.gam(model1)
summary(model1)
k.check(model1)
gam.check(model1)

linear_model <- gam(dengue_cases~temperature)
smooth_model <- gam(dengue_cases~s(temperature))

AIC(linear_model, smooth_model)

```


```{r}
library(readxl)
library(nlme)
library(gam)
library(sdm)
library(raster)
library(ncdf4)
library(mgcv)

attach(data_total)
model2 <- gam(incidence~s(population_count, k=5), family="quasipoisson")
plot.gam(model2)
summary(model2)
k.check(model2)
gam.check(model2)



cor = cor(data_total$incidence, data_total$population_count, use = "pairwise.complete.obs")
```







