---
title: "Dana_Testdokument_Tempplots"
output: html_document
date: "2023-06-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data_total
```

```{r}
#ggplot template

# Libraries
library(ggplot2)

# create data
xValue <-
yValue <- 
data <- data.frame(xValue,yValue)

# Plot
ggplot(data, aes(x=xValue, y=yValue)) +
  geom_line()

```

Temperature of amnat charoen over the years

```{r}
month_dates <- seq(as.Date("2006-01-01"), as.Date("2020-12-01"), by = "month", format = "%m")
plot(x = month_dates[c(1:180)], y = data_total$temperature[c(1:180)], type = "l",xlab="Time", ylab="Temperature", col="red")
lmodel=lm(data_total$temperature[c(1:180)] ~ month_dates[c(1:180)])
abline(lmodel,col="blue")
axis(4, seq(20,32))


```

algorithm test for 01/2006

```{r}

  tempvec_01_2006 <- c()
  for (n in c(1:77)) {
   temp_01_2006 <- data_total$temperature[1 + 180*n]
   tempvec <- c(tempvec, temp_01_2006)
    
  }
tempvec 
```

temperature of amnat charoen over the years

```{r}
month_dates <- seq(as.Date("2006-01-01"), as.Date("2020-12-01"), by = "month", format = "%m")
plot(x = month_dates[c(1:180)], y = data_total$temperature[c(1:180)], type = "l",xlab="Time", ylab="Temperature", col="red", main="Temperature in Amnat Charoen")
lmodel=lm(data_total$temperature[c(1:180)] ~ month_dates[c(1:180)])
abline(lmodel,col="blue")
axis(4, seq(20,32))


```

creating vector of mean temp for thailand for every month

```{r}

 mean_tempvec <- c()
for (month in c(1:180)) {
   tempvec <- c()
  for (district in c(1:77)) {
   temp <- data_total$temperature[month + 180*(district-1)]
   tempvec <- c(tempvec,temp)
  }
 mean_tempvec <- c(mean_tempvec,mean(tempvec))
 }
mean_tempvec
```

*linear (or other) regression of temperature*

```{r}
lmodel_thailand_temp = lm(mean_tempvec ~ month_dates[c(1:180)])

```

```         
```

#plot of temperature in thailand from 2006 to 2022

```{r}
month_dates <- seq(as.Date("2006-01-01"), as.Date("2020-12-01"), by = "month", format = "%m")
plot(x = month_dates[c(1:180)], y = mean_tempvec, type = "l",xlab="Time", ylab="Temperature", col="red")

abline(lmodel_thailand_temp,col="blue")
axis(4, seq(20,32))
```

ggplot

```{r}

# Libraries
library(ggplot2)

# create data
xValue <- month_dates
yValue <- mean_tempvec
data <- data.frame(xValue,yValue)

# Plot


ggplot() +
  geom_line(data, mapping=aes(x=xValue, y=yValue)) +
   stat_smooth(method = "lm", col = "red", mapping=aes(x=xValue, y=yValue)) +
  xlab("Years") +
  ylab("Temperature [°C]")


```

Computing minimum and maximum average temperature

```{r}
max_mean_tempvec = max(mean_tempvec)

min_mean_tempvec = min(mean_tempvec)

max_mean_tempvec
min_mean_tempvec


```

Plot minimum mean temperature of Thailand

```{r}

# Set the desired time frame
start_date_2013 <- as.Date("2013-01-01")
end_date_2014 <- as.Date("2014-12-31")

# Select values within the time frame
subset_temp_1314 <- subset(mean_tempvec, month_dates >= start_date_2013 & month_dates <= end_date_2014)

# Create a sequence of indices for the selected time frame
month_dates_1314 <- seq(as.Date("2013-01-01"), as.Date("2014-12-31"), by = "month", format = "%m")


# Plot the selected values
data_1314 = data.frame(month_dates_1314, subset_temp_1314)
ggplot(data_1314, mapping=aes(x=month_dates_1314, y=subset_temp_1314)) +
  geom_line() +
  geom_point()+
  xlab("Months") +
  ylab("Temperature [°C]") +
  ggtitle("Monthly mean temperature in Thailand 2013 - 2014")



```

Function for plotting temperature in individual timeframe

```{r}

subset_mean_temp = function(start, end, mean_tempvec, month_dates){
  # input as string
  start_date = as.Date(start)
  end_date = as.Date(end)
  
  subset_temp = subset(mean_tempvec, month_dates >= start_date & month_dates <= end_date)
  
  month_dates_year = seq(start_date, end_date, by = "month", format = "%m")
  
  data_year = data.frame(month_dates_year, subset_temp)
  
  ggplot(data_year, mapping = aes(x=month_dates_year, y=subset_temp))+
    geom_line()+
    geom_point()+
    xlab("Months") +
    ylab("Temperature [°C]") +
    ggtitle(paste("Monthly mean temperature in Thailand", start, " - ", end))
}

```

Plot timeframe of min and max temperature

```{r}
subset_mean_temp("2013-01-01", "2014-12-31", mean_tempvec, month_dates)
subset_mean_temp("2016-01-01", "2016-12-31", mean_tempvec, month_dates)
```
Probieren Inzidenz Funktion mit Vektor von Monaten 
```{r}

library(sf)
library(ggplot2)


district_sf <- st_read("gadm36_THA_shp/gadm36_THA_1.shp")

incidence_months = c(rows_2013Dec, rows_2016Apr, rows_2011Jun)

map_incidence_monthly = function(incidence_months){
  for(incidence_month in incidence_months){
 max_incidence = which.max(incidence_month$incidence)
max_incidence_area = incidence_month$Reporting_areas[max_incidence]



ggplot() +
  geom_sf(data = district_sf, aes(fill = incidence_month$incidence)) +
  scale_fill_gradient(low = "yellow", high = "red") +
  labs(fill = paste("Highest incidence =", max_incidence, "in", max_incidence_area))
  
  }
}
#input for incidence month in format: rows_YYYY_Mon



map_incidence_monthly(incidence_months)


```

