---
title: "Testdokument BioInfo Projekt"
author: "Me"
date: "2023-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("readxl")
#install.packages("ggplot2")
#install.packages("scales")
library(readxl)
```

##Einlesen der Tabellen und umwandeln der Tabellen in data frames

```{r}
library(readxl)
library(dplyr)

Endung <- "_dengue_extracted.xlsx"
Anfang <- "Dengue"

for (i in 2006:2020) {
  
  Dateiname <- paste0(i, Endung) #erstellen das Dateinamens
  Name_df <- paste0(Anfang, i) 
  assign(Name_df, read_excel(Dateiname)) #erstellen der Variable und zuweisen der Excel-Tabelle
  
  df_dengue <- get(Name_df)
  vec_data_names <- vector("character")
  
  vec_data_names <- sapply(df_dengue$Reporting_areas, function(g) {
    paste(g[1], i)
  })
  data_names <- matrix(vec_data_names,nrow = 77, ncol = 1)
  df_dengue <- cbind(data_names, df_dengue)
  assign(Name_df,df_dengue)
}
total_years = data.frame()
for (i in 2006:2020){
  a = get(paste0("Dengue",i))
  total_years = rbind(total_years, a)
}
```

#Rows als Zeitachse mit Monaten

```{r}
m = 1
data_total = data.frame(col1 = character(0), col2 = character(0), col3 = numeric(0), month = character(0))
colnames(data_total) = c("data_names", "Reporting_areas", "dengue_cases", "month")
sorted_total_years = total_years[order(total_years$"Reporting_areas"),]
for (k in 1:1155){
  for (j in 3:14){
    square <- sorted_total_years[k, c(1, 2, j)]
    month = colnames(square[3])
    square <- cbind(square,month)
    colnames(square) = c("data_names", "Reporting_areas", "dengue_cases", "month")
    data_total <- rbind(data_total, square)
    colnames(data_total) = c("data_names", "Reporting_areas", "dengue_cases", "month")
    data_total[m,1] = paste(data_total[m,1], month[1])
    m = m + 1
  }
}
```

#adding temperature, longitude and latitude column

```{r}
m = 1
temp_data = read_excel("era5_data_2006_2020_thailand_monmean.xlsx")
temp_data = as.data.frame(temp_data)
temp_data_sorted = data.frame( col1 = numeric(0), Longitude = numeric(0), Latitude = numeric(0))
for (i in 1:77) {
  for (j in 5:184) {
    temp_data_sorted[m,1] = temp_data[i,j]
    temp_data_sorted[m,2] = temp_data[i,3]
    temp_data_sorted[m,3] = temp_data[i,4]
    m = m + 1
  }
}
data_total = cbind(data_total, temp_data_sorted)
colnames(data_total)[5] = "temperature"
```

#Einlesen Population data

```{r}
library(readxl)

Endung <- "_population.xlsx"
#empty dataframe for the data
total_population = data.frame(Reporting_areas = character(), population_count = numeric())

# stack years of 2006-2011 below each other
for (i in 2006:2011) {
  population_df  = data.frame()
  Dateiname <- paste0(i, Endung) #erstellen das Dateinamens
  population_df = read_excel(Dateiname) #erstellen der Variable und zuweisen der Excel-Tabelle
  vec_years = rep(i, times = 77)
  population_df = cbind(population_df,vec_years)
  names(population_df) = c("Reporting_areas", "population_count", "year")
  # multiply every row 12 times for each month
  repeat_count <- 12
  population_df_replicated <- population_df[rep(row.names(population_df), each = repeat_count), ]
  row.names(population_df_replicated) <- NULL
  
  #put all the years together
  total_population = rbind(total_population, population_df_replicated)
  
}

# stack years of 2012-2020 below each other and add to data_total
population_df = read_excel("2012-2020_population.xlsx")
colnames(population_df) = c("Reporting_areas", "population_count", "population_count", "population_count", "population_count", "population_count","population_count","population_count","population_count","population_count")
pop_placeholder = data.frame(Reporting_areas = character(), population_count = numeric(), year = numeric())
col_num = 2
for (i in 2012:2020) {
  year = rep(i, times = 77)
  pop_placeholder = rbind(pop_placeholder, cbind(population_df[,c(1,col_num)],year))
  names(pop_placeholder) = c("Reporting_areas", "population_count", "year")
  col_num = col_num + 1
}

# multiply every row 12 times for each month
repeat_count <- 12
population_df_replicated <- pop_placeholder[rep(row.names(pop_placeholder), each = repeat_count), ]
row.names(population_df_replicated) <- NULL

# combine all population data in each 
total_population = rbind(total_population, population_df_replicated)
sorted_total_population = total_population[order(total_population$"Reporting_areas"),]

data_total = cbind(data_total,sorted_total_population[,c(2,3)])
row.names(data_total) = NULL
```

# adding time column

```{r}
first_date = as.Date("2006-01-01")
last_date = as.Date("2020-12-31")

time_column_short = seq(first_date, last_date, by = "month")
time_column_short = as.Date(time_column_short)
time_column = rep(time_column_short, times = 77)

data_total = cbind(data_total, time_column)
#data_total = subset(data_total, select = -time_column)
```

#Detect missing values

```{r}
detect_na <- function(x) {
  is_na <- x %in% c("N/A", "na", "NA", "n/a")
  replace(x, is_na, NA)
}
data_total <- apply(data_total, 2, detect_na)
rmv.rows = apply(data_total,1,function(x){sum(is.na(x))})
i.missing = which(rmv.rows >0)
```

#Inzindenz

```{r}
data_total = as.data.frame(data_total)
data_total$dengue_cases = as.integer(data_total$dengue_cases)
data_total$population_count = as.integer(data_total$population_count)
dimensions = dim(data_total)
R= dimensions[1]

for(Zeile in 1:R){
  Population <- data_total$population_count[Zeile]
  Infection <- data_total$dengue_cases[Zeile]
  Incidence = (Infection/Population)*100000
  data_total$incidence[Zeile] <- Incidence
}
```

#create clean data frame

```{r}
clean_data_total = as.data.frame(data_total[-i.missing,])
num_col = c("temperature", "Longitude", "Latitude", "incidence")
int_col = c("dengue_cases","population_count","year")
date_col = c("time_column")
clean_data_total[int_col] = lapply(clean_data_total[int_col], as.integer)
clean_data_total[num_col] = lapply(clean_data_total[num_col], as.numeric)
clean_data_total[date_col] = lapply(clean_data_total[date_col], as.Date)
summary(clean_data_total)
```

#create data frame for monsoon season

```{r}
vec_month_monsoon = c("Jun", "Jul", "Aug", "Sep")
data_monsoon = data_total[data_total$month %in% vec_month_monsoon,]

rmv.rows.m = apply(data_monsoon,1,function(x){sum(is.na(x))})
i.missing.m = which(rmv.rows.m >0)
clean_data_monsoon = as.data.frame(data_monsoon[-i.missing.m,])

```

#struture data frame and convert into right datatypes

```{r}
data_total = as.data.frame(data_total)
num_col = c("temperature", "Longitude", "Latitude", "incidence")
int_col = c("dengue_cases","population_count","year")
date_col = c("time_column")
data_total[int_col] = lapply(data_total[int_col], as.integer)
data_total[num_col] = lapply(data_total[num_col], as.numeric)
data_total[date_col] = lapply(data_total[date_col], as.Date)

data_total = subset(data_total, select = c("data_names", "Reporting_areas", "year", "month", "time_column", "dengue_cases", "population_count", "temperature", "Longitude", "Latitude", "incidence"))

save(data_total, file = "data_total.RData")
summary(data_total)
```

#max, min, sd

```{r}
apply(data_total[,c(num_col, int_col)], 2, function(x){max(x,na.rm=TRUE)})
apply(data_total[,c(num_col, int_col)], 2, function(x){min(x,na.rm=TRUE)})
apply(data_total[,c(num_col, int_col)], 2, function(x){mean(x,na.rm=TRUE)})
apply(data_total[,c(num_col, int_col)], 2, function(x){median(x,na.rm=TRUE)})
apply(data_total[,c(num_col, int_col)], 2, function(x){sd(x,na.rm=TRUE)})
```

#mean of dengue_cases and temperature of whole thailand per month

```{r}
library(ggplot2)
library(scales)

#dengue cases of whole thailand per month
f_mean_dengue = function(){
  mean_dengue <- data.frame(dengue_cases = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(dengue_cases = mean(month$dengue_cases, na.rm = TRUE), time_column = month$time_column[1])
    mean_dengue = rbind(mean_dengue, m_month)
  }
  return(mean_dengue)
}


f_mean_temp = function(){
  mean_temp <- data.frame(temperature = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(temperature = mean(month$temperature, na.rm = TRUE), time_column = month$time_column[1])
    mean_temp = rbind(mean_temp, m_month)
  }
  return(mean_temp)
}

plot1 = ggplot2::ggplot() +
  geom_line(data = f_mean_dengue(), aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases") +
  scale_x_continuous(breaks = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "2 year"))

plot2 = ggplot2::ggplot() +
  geom_line(data = f_mean_temp(), aes(x = time_column, y = temperature), color = "red") +
  labs(x = "time", y = "temperature") +
  scale_x_continuous(breaks = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "2 year"))


gridExtra::grid.arrange(plot1, plot2, nrow = 2)
```

#mean and median comparison

```{r}
library(ggplot2)

f_median_dengue = function(){
  median_dengue <- data.frame(dengue_cases = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(dengue_cases = median(month$dengue_case, na.rm = TRUE), time_column = month$time_column[1])
    median_dengue = rbind(median_dengue, m_month)
  }
  return(median_dengue)
}

f_median_temp = function(){
  median_temp <- data.frame(temperature = numeric(), time_column = as.Date(character()))
  months = seq(from = as.Date("2006-01-01"), to = as.Date("2020-12-01"), by = "month")
  for (i in months) {
    month = data_total[data_total$time_column == i,]
    m_month = data.frame(temperature = median(month$temperature, na.rm = TRUE), time_column = month$time_column[1])
    median_temp = rbind(median_temp, m_month)
  }
  return(median_temp)
}
#compute difference between mean and median
dif_mm_dengue = f_mean_dengue()["dengue_cases"]-f_median_dengue()["dengue_cases"]
dif_mm_dengue = cbind(dif_mm_dengue,f_mean_dengue()["time_column"])

dif_mm_temp = f_mean_temp()["temperature"]-f_median_temp()["temperature"]
dif_mm_temp = cbind(dif_mm_temp,f_mean_temp()["time_column"])

plot_mean_temp = ggplot2::ggplot() +
  geom_line(data = f_mean_temp(), aes(x = time_column, y = temperature)) +
  labs(x = "time", y = "temperature", title = "Mean temperature")

plot_mean_dengue = ggplot2::ggplot() +
  geom_line(data = f_mean_dengue(), aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases", title = "Mean dengue cases")

plot_median_temp = ggplot2::ggplot() +
  geom_line(data = f_median_temp(), aes(x = time_column, y = temperature)) +
  labs(x = "time", y = "temperature", title = "Median temperature")

plot_median_dengue = ggplot2::ggplot() +
  geom_line(data = f_median_dengue(), aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases", title = "Median dengue cases")

plot_dif_dengue = ggplot2::ggplot() +
  geom_line(data = dif_mm_dengue, aes(x = time_column, y = dengue_cases)) +
  labs(x = "time", y = "dengue-cases", title = "Difference between mean and median")

plot_dif_temp = ggplot2::ggplot() +
  geom_line(data = dif_mm_temp, aes(x = time_column, y = temperature)) +
  labs(x = "time", y = "temperature", title = "Difference between mean and median")

gridExtra::grid.arrange(plot_mean_dengue, plot_mean_temp, plot_median_dengue, plot_median_temp,plot_dif_dengue,plot_dif_temp, nrow = 3)
```

#mean of dengue_cases and temperature of whole thailand per year

```{r}
f_mean_dengue.y.t = function(){
  mean_dengue.y.t <- data.frame(dengue_cases = numeric(), time_column = as.Date(character()))
  for (i in c(2006:2020)) {
    month = data_total[data_total$year == i,]
    m_month = data.frame(dengue_cases = mean(month$dengue_cases, na.rm = TRUE), time_column = month$time_column[1])
    mean_dengue.y.t = rbind(mean_dengue.y.t, m_month)
  }
  return(mean_dengue.y.t)
}
mean_dengue.y.t = as.data.frame(f_mean_dengue.y.t())

f_mean_temp.y.t = function(){
  mean_temp.y.t <- data.frame(temperature = numeric(), time_column = as.Date(character()))
  for (i in c(2006:2020)) {
    month = data_total[data_total$year == i,]
    m_month = data.frame(temperature = mean(month$temperature, na.rm = TRUE), time_column = month$time_column[1])
    mean_temp.y.t = rbind(mean_temp.y.t, m_month)
  }
  return(mean_temp.y.t)
}
mean_temp.y.t = as.data.frame(f_mean_temp.y.t())

```

#mean of incidence and temperature of each provinze per year

```{r}
f_mean_inc.y.p = function(period, areas){
  mean_inc.y.p <- data.frame(incidence = numeric(), time_column = as.Date(character()), Reporting_areas = character())
  for (i in period){
    year = data_total[data_total$year == i,]
    for (p in areas) {
      provinz = year[year$Reporting_areas == p,]
      m_y_p = data.frame(incidence = mean(provinz$incidence, na.rm = TRUE), time_column = provinz$time_column[1], Reporting_areas = provinz$Reporting_areas[1])
      mean_inc.y.p = rbind(mean_inc.y.p, m_y_p)
    }
  }
  return(mean_inc.y.p)
}

f_mean_temp.y.p = function(period, areas){
  mean_temp.y.p <- data.frame(temperature = numeric(), time_column = as.Date(character()), Reporting_areas = character())
  for (i in period){
    year = data_total[data_total$year == i,]
    for (p in areas) {
      provinz = year[year$Reporting_areas == p,]
      m_y_p = data.frame(temperature = mean(provinz$incidence, na.rm = TRUE), time_column = provinz$time_column[1], Reporting_areas = provinz$Reporting_areas[1])
      mean_temp.y.p = rbind(mean_temp.y.p, m_y_p)
    }
  }
  return(mean_temp.y.p)
}
```

#Plot für Bangkok

```{r}
dengue_plot_cases = c()
dengue_plot_time = c()

dengue_plot_cases <- data_total$dengue_cases[data_total$Reporting_areas == "Bangkok"]
dengue_plot_time <- data_total$time_column[data_total$Reporting_areas == "Bangkok"]

plot(dengue_plot_time,dengue_plot_cases, type = "o")
```

#distribution of incidence

```{r}
q.dengue = quantile(data_total$incidence, probs = c(0.01,0.1,0.25,0.5,0.75,0.9,0.99), na.rm = TRUE)
hist(data_total$incidence,  breaks = 200, main = "Distribution of incidence", xlab = "incidence", freq = TRUE);abline(v=q.dengue,lty=3,lwd=2,col='red')

q.temp = quantile(data_total$temperature, probs = c(0.01,0.1,0.25,0.5,0.75,0.9,0.99), na.rm = TRUE)
hist(data_total$temperature,  breaks = 200, main = "Distribution of temperature", xlab = "temperature", freq = TRUE);abline(v=q.temp,lty=3,lwd=2,col='red')
```

#correlation between temperature and incidence

```{r}
library(npreg)
library(ggplot2)

smoothed_curve <- ss(clean_data_total$temperature, clean_data_total$incidence, nknots = 10)

plot(clean_data_total$temperature, clean_data_total$incidence, pch = 20,
     ylab = "incidence",
     xlab = "temperature",
  abline(v = mean(clean_data_total$temperature), col = "red"))

plot(smoothed_curve, col = "blue")

plot(rank(data_total$temperature), rank(data_total$incidence), pch = 20,
     ylab = "incidence",
     xlab = "temperature",
  abline(v = mean(data_total$temperature), col = "red"))

qqplot(data_total$temperature, data_total$incidence)

#pearson correlation y.t.
cor(mean_dengue.y.t$dengue_cases, mean_temp.y.t$temperature)
#spearman correlation y.t.
cor(mean_dengue.y.t$dengue_cases, mean_temp.y.t$temperature, method = "spearman")

#pearson correlation m.p.
cor(clean_data_total$dengue_cases, clean_data_total$temperature)
#spearman correlation m.p.
cor(clean_data_total$dengue_cases, clean_data_total$temperature, method = "spearman")


pheatmap::pheatmap(cor(clean_data_total[c("dengue_cases", "population_count", "temperature", "incidence")]))

```

#monsoon season smoothed curve incidence temperature

```{r}
library(npreg)
library(ggplot2)

smoothed_curve_m <- ss(clean_data_monsoon$temperature, clean_data_monsoon$incidence, nknots = 10)
plot(clean_data_monsoon$temperature, clean_data_monsoon$incidence, pch = 20,
     ylab = "incidence",
     xlab = "temperature",
  abline(v = mean(clean_data_monsoon$temperature), col = "red"))
plot(smoothed_curve, col = "blue")

#only for quantile 5% to 95%
lower_limit = quantile(clean_data_monsoon$temperature, probs = 0.05, na.rm = T)
upper_limit = quantile(clean_data_monsoon$temperature, probs = 0.95, na.rm = T)

clean_data_monsoon_5.95 = subset(clean_data_monsoon, temperature >= lower_limit & temperature <= upper_limit)

smoothed_curve_m_5.95 <- ss(clean_data_monsoon_5.95$temperature, clean_data_monsoon_5.95$incidence, nknots = 10)
plot(clean_data_monsoon_5.95$temperature, clean_data_monsoon_5.95$incidence, pch = 20,
     ylab = "incidence",
     xlab = "temperature",
  abline(v = mean(clean_data_monsoon_5.95$temperature), col = "red"))
plot(smoothed_curve_m_5.95, col = "blue")

```

#correlation between different areas

```{r}
#mean of temp and incidence over all years per area
t_data_total = t(data_total)
r_areas = unique(data_total$Reporting_areas)
m_inc_temp.y.r = data.frame(Reporting_areas = character(), incidence = numeric(), temperature = numeric())
for (i in r_areas){
  temp_inc = apply(data_total[which(clean_data_total$Reporting_areas == i),c("incidence", "temperature")], 2,function(g){mean(g, na.rm = TRUE)})
  temp_inc_df = data.frame(Reporting_areas = i, incidence = temp_inc["incidence"], temperature = temp_inc["temperature"])
  m_inc_temp.y.r = rbind(m_inc_temp.y.r, temp_inc_df)
  rownames(m_inc_temp.y.r) <- 1:nrow(m_inc_temp.y.r)
}

```

#k-means clustering

```{r}
#elbow method
km = kmeans(m_inc_temp.y.r["incidence"], centers = 2, nstart = 10)
km$tot.withinss

wss = sapply(2:7,function(k) { 
  kmeans(m_inc_temp.y.r["incidence"], centers = k)$tot.withinss
})
plot(2:7,wss,type='b',pch=19,xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")


#silhoutte method
library(cluster)
#library (NbClust)
#library (clustertend)
library (factoextra)
#library (fpc)
#library (clValid)
library(ggplot2)

#find optimal number of clusters by incidence
D_inc = dist(m_inc_temp.y.r["incidence"])
km_inc = kmeans(m_inc_temp.y.r["incidence"], centers = 4, nstart = 10)
s_inc = silhouette(km_inc$cluster,D_inc)
plot(s_inc)
fviz_nbclust(m_inc_temp.y.r["incidence"], pam, method = "silhouette")+ theme_classic()

#find optimal number of clusters by temperature
D_temp = dist(m_inc_temp.y.r["temperature"])
km_temp = kmeans(m_inc_temp.y.r["temperature"], centers = 4, nstart = 10)
s_temp = silhouette(km_temp$cluster,D_temp)
plot(s_temp)
fviz_nbclust(m_inc_temp.y.r["temperature"], pam, method = "silhouette")+ theme_classic()

#find optimal number of clusters by temperature and incidence
D_inc_temp = dist(scale(m_inc_temp.y.r[,c("incidence","temperature")]))
km_inc_temp = kmeans(scale(m_inc_temp.y.r[,c("incidence","temperature")]), centers = 3)
s_inc_temp = silhouette(km_inc_temp$cluster,D_inc_temp)
plot(s_inc_temp)
fviz_nbclust(scale(m_inc_temp.y.r[,c("incidence","temperature")]), pam, method = "silhouette")+ theme_classic()

m_inc_temp_cluster.y.r = data.frame(m_inc_temp.y.r, km_inc_temp$cluster)

plot1 = ggplot(m_inc_temp.y.r, aes(x = temperature, y = incidence, color = as.factor(km_inc$cluster))) +
  geom_point() +
  labs(title = "Partitioning Clustering Plot by incidence") +
  theme_classic()

plot2 = ggplot(m_inc_temp.y.r, aes(x = temperature, y = incidence, color = as.factor(km_temp$cluster))) +
  geom_point() +
  labs(title = "Partitioning Clustering Plot by temperature") +
  theme_classic()
plot2

plot3 = ggplot(m_inc_temp.y.r, aes(x = temperature, y = incidence, color = as.factor(km_inc_temp$cluster))) +
  geom_point() +
  labs(title = "Partitioning Clustering Plot") +
  theme_classic()
plot3
gridExtra::grid.arrange(plot1, plot2, plot3, ncol = 1)
```

#probieren mit time series Objekten

```{r}
my_ts <- ts(data_total$dengue_cases, start = c(data_total$time_column[1]), frequency = 12)
head(my_ts)
TSstudio::ts_plot(my_ts)
```

#GAM

```{r}
library(readxl)
library(nlme)
library(gam)
library(sdm)
library(raster)
library(ncdf4)
library(mgcv)

attach(data_total)
model1 = gam(incidence~s(temperature, k = 16), family = "quasipoisson", na.rm = T)
plot.gam(model1)
summary(model1) #edf: → a value close to 1 tends to be close to a linear term → a higher value means that the spline is more wiggly (non-linear). 
#F-Wert und der p-Wert geben die statistische Signifikanz der glatten Funktion an. In diesem Fall ist die glatte Funktion von "temperature" hoch signifikant (p-Wert < 2e-16), was darauf hindeutet, dass sie einen signifikanten Einfluss auf die Vorhersage der "incidence" hat.Der angegebene R-squared-Wert (R-sq.(adj)) von 0.0407 gibt an, wie gut das Modell die Varianz der abhängigen Variable erklärt. In diesem Fall erklärt das Modell etwa 4,07% der Varianz in den Daten. Ein niedriger GCV-Wert deutet auf eine gute Anpassung des Modells hin.
k.check(model1)
gam.check(model1)


best_aic = Inf
best_model = NULL
k_best = 0
for (i in 3:50){
      # Calculate AIC
      smooth_model = gam(incidence~s(temperature, k = i))
      aic = AIC(smooth_model)
      # Check if the current model has lower AIC than the best model so far
      if (aic < best_aic) {
        best_aic = aic
        best_model = model
        k_best = i
      }
}
best_aic
k_best

smooth_model = gam(incidence~s(temperature, k = 16))
linear_model = gam(incidence~temperature, k = 16)
AIC(smooth_model, linear_model)

detach(data_total)
```

# GAM forecast

```{r, fig.cap="GAM prediction", fig.ext="png", knitr::include_graphics(plot_gam_pred)}
library(raster)
library(gam)
library(sdm)
library(readxl)
library(mgcv)
library(ncdf4)

## lower_limit = quantile(data_monsoon$temperature, probs = 0.05, na.rm = T)
#upper_limit = quantile(data_monsoon$temperature, probs = 0.95, na.rm = T)

#data_monsoon_5.95 = subset(data_monsoon, temperature >= lower_limit & temperature <= upper_limit)

attach(data_total)

model2 <- gam(incidence~s(temperature, k = 16), family = "quasipoisson")
tas <- stack("tas_SEA22_MPI_rcp85_2021-2040_grid_subc_daymean_monmean_timmean.nc")
#tas_test = crop(tas, my.area)
names(tas) <- "temperature"
p1 <- raster::predict(tas, model2, type = "response")

thailand <- raster::getData('GADM', country='THA', level=1)
my.area <- extent(thailand)
my.p1 <- crop(p1, my.area)
my.p1.mask <- mask(my.p1, thailand)
plot(my.p1.mask, legend.width=1, legend.shrink=0.75, axes= FALSE, box=FALSE) + title("Predicted values: Influence of temperature on dengue cases for 2021-2040", line = +1) + plot(thailand, add=TRUE)

detach(data_total)
```

#Plotting der Inzidenz eines einzelnen Monats auf die Karte von Thailand

```{r}
library(sf)
library(ggplot2)


district_sf <- st_read("gadm36_THA_shp/gadm36_THA_1.shp")

#Inzidenz vom April 2006 auf die karte geplottet

district_sf$incidence2006Aug<- rows_2006Aug$incidence

ggplot() +
  geom_sf(data = district_sf, aes(fill = incidence2006Aug)) +
  scale_fill_gradient(low = "yellow", high = "red") + labs(fill = "Incidence in Aug 2006")
```

#Funktion für Inzidenz plot der Monate

```{r}

library(sf)
library(ggplot2)


district_sf <- st_read("gadm36_THA_shp/gadm36_THA_1.shp")


map_incidence_monthly = function(incidence_month){
  
 max_incidence = which.max(incidence_month$incidence)
max_incidence_area = incidence_month$Reporting_areas[max_incidence]



ggplot() +
  geom_sf(data = district_sf, aes(fill = incidence_month$incidence)) +
  scale_fill_gradient(low = "white", high = "green") +
  labs(fill = paste("Highest incidence =", max_incidence, "in", max_incidence_area))
  
  
}
#input for incidence month in format: rows_YYYY_Mon

map_temp_monthly = function(temp_month){
  
max_temp = which.max(temp_month$temperature)
max_temp_area = temp_month$Reporting_areas[max_temp]



ggplot() +
  geom_sf(data = district_sf, aes(fill = temp_month$temperature)) +
  scale_fill_gradient(low = "white", high = "green") +
  labs(fill = paste("Highest temperature =", max_temp, "in", max_temp_area))
  
  
}
map_temp_monthly(rows_2006)

gridExtra::grid.arrange(map_temp_monthly(f_mean_temp.y.p(period = c(2020), areas = r_areas)),map_incidence_monthly(f_mean_inc.y.p(period = c(2020), areas = r_areas)))

```
